---
title: "Final project for practical machine learning"
output: html_document
---

To build a machine learning model for predicting the classe of individuals using their motion data, I first organized and cleaned the training data sets. Then I further split the orginal training sets into training and testing data sets. I used random forest tree method to build the prediction model and tested on the testing data sets. With the success in predicing classe values in testing data sets, I used the model to predict classe values using the given motion data.

### Step 1. Organize and clean the training data
The datasets containing lots of variables with NA values and empty values, so these variables are removed to reduce the computation loads. Then the name variable is also removed. 

```{r}
setwd("~/desktop/data science/project 4")
fitdata1=read.csv("pml-training.csv")
fitdata2=read.csv("pml-testing.csv")
fitdata1[fitdata1==""]=NA
fitdata2[fitdata2==""]=NA
cleandata1=fitdata1[,colSums(is.na(fitdata1))<100]
cleandata2=fitdata2[,colSums(is.na(fitdata2))<5]
cleandata1=cleandata1[c(-1,-2)]
```
### Step 2. Organize the orginal training datasets into testing and training datasets

```{r}
set.seed(1233)
library(caret)
intrain=createDataPartition(cleandata1$class,p=0.6,list=FALSE)
train=cleandata1[intrain,]
test=cleandata1[-intrain,]
```
### Step 3. Build the prediction model using random forest tree method
Typically having 50-100 trees will give good results with reasonable amount of computation time

```{r}

modfit=train(classe~.,data=train,method="rf",ntree=50)
modfit
```

### Step 4. Validate the prediction model using testing data sets

```{r}
pred=predict(modfit,test)
confusionMatrix(pred,test$classe)
```
---
Validation Results:
Confusion Matrix and Statistics

          Reference
Prediction    A    B    C    D    E
         A 2232    3    0    0    0
         B    0 1515    1    0    0
         C    0    0 1367    5    0
         D    0    0    0 1280    0
         E    0    0    0    1 1442

Overall Statistics
                                          
               Accuracy : 0.9987          
                 95% CI : (0.9977, 0.9994)
    No Information Rate : 0.2845          
    P-Value [Acc > NIR] : < 2.2e-16       
                                          
                  Kappa : 0.9984          
 Mcnemar's Test P-Value : NA              

Statistics by Class:

                     Class: A Class: B Class: C Class: D Class: E
Sensitivity            1.0000   0.9980   0.9993   0.9953   1.0000
Specificity            0.9995   0.9998   0.9992   1.0000   0.9998
Pos Pred Value         0.9987   0.9993   0.9964   1.0000   0.9993
Neg Pred Value         1.0000   0.9995   0.9998   0.9991   1.0000
Prevalence             0.2845   0.1935   0.1744   0.1639   0.1838
Detection Rate         0.2845   0.1931   0.1742   0.1631   0.1838
Detection Prevalence   0.2849   0.1932   0.1749   0.1631   0.1839
Balanced Accuracy      0.9997   0.9989   0.9992   0.9977   0.9999
---
The confusion matrix shows the model has great sensitivity.

### Step 5. Use the built model to predict casse for given data sets

```{r}
newclasse=predict(modfit,cleandata2)
print(newclasse)

```

---
Predictions: [1] B A B A A E D B A A B C B A E E A B B B
Levels: A B C D E
100% accuracy using my model
---
